---
title: "Efficient Portfolio"
output:
  html_document: default
  html_notebook: default
---


# Calculating an efficient portfolio with NEPC data

We will calculate the minimum variance portfolio with target return of 8%.  

```{r, echo=FALSE}
source('portfolio_noshorts.r.txt', encoding='UTF-8')
require(knitr)
# function to convert from geom return to arith return
# geo is the geometric return (may be a vector)
# vol is the volatility of the asset (may be a vecor)
# n is the number of years
geom_to_arith=function(geo,vol,n) {
  var=vol^2
  factor=(1+var/((1+geo)^2))^((1-n)/(2*n))
  arith=-1+(1+geo)/factor
  return(arith)
}
# function to convert from arith return to geom return
# arith is the arithmetic return (may be a vector)
# vol is the volatility of the asset (may be a vecor)
# n is the number of years
arith_to_geom=function(arith,vol,n) {
  var=vol^2
  factor=(1+var/((1+arith)^2))^((1-n)/(2*n))
  geo=-1+(1+arith)*factor
  return(geo)
}
```
```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,8,6.5)/100
volatility=c(6.03,17.5,21,21,28,13,9.29,15)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,70,60,40,
                       0,90,100,60,65,75,65,40,
                       0,70,60,100,70,60,75,35,
                       0,60,65,70,100,45,80,30,
                       0,70,75,60,45,100,65,50,
                       0,60,65,75,80,65,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

ans=efficient.portfolio(er=geom_returns,
                       cov.mat=covariance,
                       target.return=.078,
                       shorts=FALSE)
port_arith_return=arith_returns%*%ans$weights
port_geom_return=arith_to_geom(port_arith_return,ans$sd,n)
```

The weights for the portfolio are `r kable(ans$weights,digits=2)`
The portfolio has a standard deviation of `r round(ans$sd*100,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%    

Since the portfolio risk are allocated entirely to private equity and private credit, this suggests the assumptions in not in line with capital markets. 

We will increase the volatility of private equity to 21% (equal to small cap stocks).  We will increase the volatility of private debt to 13%, similar to high yield.  We reduce the volatility of of large cap US stocks to 15%, more in line with recent history.  We will reduce the volatility of real estate to 12% which is consistent with the desmoothing study I did a few years ago.  We leave the returns alone for now.

```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,8,6.5)/100
volatility=c(6.03,15,21,21,28,21,13,12)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,70,60,40,
                       0,90,100,60,65,75,65,40,
                       0,70,60,100,70,60,75,35,
                       0,60,65,70,100,45,80,30,
                       0,70,75,60,45,100,65,50,
                       0,60,65,75,80,65,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

ans=efficient.portfolio(er=geom_returns,
                       cov.mat=covariance,
                       target.return=.078,
                       shorts=FALSE)
port_arith_return=arith_returns%*%ans$weights
port_geom_return=arith_to_geom(port_arith_return,ans$sd,n)
```

The weights for the portfolio are `r kable(ans$weights,digits=2)`
The portfolio has a standard deviation of `r round(ans$sd*100,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%    


Let's reduce the assumed return for private debt to 7.25%, which is lower than large cap equity returns.  We will increase the real estate return 7.25% which is more compatible with the strategy we are following.


```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,7.25,7.25)/100
volatility=c(6.03,15,21,21,28,21,13,12)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,70,60,40,
                       0,90,100,60,65,75,65,40,
                       0,70,60,100,70,60,75,35,
                       0,60,65,70,100,45,80,30,
                       0,70,75,60,45,100,65,50,
                       0,60,65,75,80,65,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

ans=efficient.portfolio(er=geom_returns,
                       cov.mat=covariance,
                       target.return=.078,
                       shorts=FALSE)
port_arith_return=arith_returns%*%ans$weights
port_geom_return=arith_to_geom(port_arith_return,ans$sd,n)
```

The weights for the portfolio are `r kable(ans$weights,digits=2)`
The portfolio has a standard deviation of `r round(ans$sd*100,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%    

Now let's increase the correlation of private equity to the rest of equities to 90% correlated with large, 100% correlated with small, 75% correlated with EAFE,60% correlated with EM, 75% correlated with private credit.



```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,7.25,7.25)/100
volatility=c(6.03,15,21,21,28,21,13,12)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,90,60,40,
                       0,90,100,60,65,90,65,40,
                       0,70,60,100,70,75,75,35,
                       0,60,65,70,100,60,80,30,
                       0,90,90,75,60,100,75,50,
                       0,60,65,75,80,75,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

ans=efficient.portfolio(er=geom_returns,
                       cov.mat=covariance,
                       target.return=.078,
                       shorts=FALSE)
port_arith_return=arith_returns%*%ans$weights
port_geom_return=arith_to_geom(port_arith_return,ans$sd,n)
```

The weights for the portfolio are `r kable(ans$weights,digits=2)`
The portfolio has a standard deviation of `r round(ans$sd*100,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%    

Now make private credit more correlated with equities.  Not logical that it is more correlated with EAFE than US.  Make 75 correlated with equities across the board.


```{r}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.5,7.75,7.75,9.5,9.5,7.25,7.25)/100
volatility=c(6.03,15,21,21,28,21,13,12)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,90,75,40,
                       0,90,100,60,65,90,75,40,
                       0,70,60,100,70,75,75,35,
                       0,60,65,70,100,60,75,30,
                       0,90,90,75,60,100,75,50,
                       0,75,75,75,75,75,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

ans=efficient.portfolio(er=geom_returns,
                       cov.mat=covariance,
                       target.return=.078,
                       shorts=FALSE)
port_arith_return=arith_returns%*%ans$weights
port_geom_return=arith_to_geom(port_arith_return,ans$sd,n)
```

The weights for the portfolio are `r kable(ans$weights,digits=2)`
The portfolio has a standard deviation of `r round(ans$sd*100,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%  

Let's switch to an optimizer that allows us to place constraints and run with the following constraints.  Core Bonds, large cap equity and EAFE have a minimum of 10%.  All other assets have a maximum of 20%, with the exception of private equity which is limited to 10% and EM which is limited to 5%.  I limit the maximum combined equity position to 60%.

```{r, echo=FALSE, warning=FALSE, message=FALSE}

library(PortfolioAnalytics)
require(ROI)
require(ROI.plugin.glpk)
require(ROI.plugin.quadprog)

my.custom.moments <- function(R, portfolio){
  out <- list()
  out$mu <- geom_returns
  out$sigma <- covariance
  out
}
```


```{r, warning=FALSE, message=FALSE}
asset_names=c("BbgAgg","LrgUSEq","SmlUSEq","EAFE","EM","PE","PrivCredit","RE")
geom_returns=c(4,7.75,8.25,7.75,8.25,8.75,7.75,7.75)/100
volatility=c(6.03,15,21,21,28,21,13,12)/100
correlation=matrix(nrow=8,data=
                     c(100,0,0,0,0,0,0,10,
                       0,100,90,70,60,90,75,40,
                       0,90,100,60,65,90,75,40,
                       0,70,60,100,70,75,75,35,
                       0,60,65,70,100,60,75,30,
                       0,90,90,75,60,100,75,50,
                       0,75,75,75,75,75,100,40,
                       10,40,40,35,30,50,40,100))/100
names(geom_returns)=names(volatility)=asset_names
rownames(correlation)=asset_names
colnames(correlation)=asset_names
n=30
arith_returns=geom_to_arith(geom_returns,volatility,n)
Dsqrt=0*correlation
diag(Dsqrt)=volatility
covariance=Dsqrt%*%correlation%*%Dsqrt

# portfolio specification
R=matrix(geom_returns,byrow=TRUE,nrow=100,ncol=8)
colnames(R)=asset_names
R=xts(R,Sys.Date()+0:99)
p <- portfolio.spec(assets = names(geom_returns))
# full investment constraint
p <- add.constraint(portfolio = p, type = "weight_sum", 
                    min_sum = 0.9999, max_sum = 1.0001)
# asset constraints
p <- add.constraint(portfolio = p, type = "box", 
                    min = c(.1,.1,0,.1,0,0,.1,.1), 
                    max = c(.3,.3,.1,.3,.1,.1,.2,.2))
p <- add.constraint(portfolio=p, type="group",
                    groups=list(groupA=c(2,3,4,5,6)),
                    group_min=.2,
                    group_max=.60)
p <- add.constraint(portfolio = p, type="return", return_target=.075)
p <- add.objective(portfolio = p, type = "return", name = "mean", multiplier = 0)
p.minvol <- add.objective(portfolio=p,type="risk",name="var")

opt.min.qp <- optimize.portfolio(R, portfolio=p.minvol, momentFUN=my.custom.moments,
                                optimize_method="ROI",trace=TRUE)

port_arith_return=arith_returns%*%opt.min.qp$weights
port_geom_return=arith_to_geom(port_arith_return,opt.min.qp$opt_values$StdDev,n)
```



The weights for the portfolio are `r kable(opt.min.qp$weights,digits=2)`
The portfolio has a standard deviation of `r round(100*opt.min.qp$opt_values$StdDev,2)`%    
The portfolio has an expected geometric return of `r round(port_geom_return*100,2)`%  
